---
title: "Tesi"
author: "Pedron Matteo"
output:
  html_document:
    toc: true
    number_sections: false
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, message=FALSE}
rm(list=ls())
gc()

library(dplyr)
library(MASS)
library(spd) 
library(sn)             # univariate / multivariate skew-t distribution
library(CVXR)

# basic finance
library(xts)   
library(quantmod)   
library(pob)            # book package with financial data
library(stochvol)       # to compute volatility envelope
library(tseries)        # jarque.bera.test
library(rugarch)
library(forecast)       # auto.arima

# plotting
library(ggplot2)
library(viridisLite)
library(gridExtra)      # for grid.arrange
library(reshape2)    
library(patchwork)      # for combining plots
library(scales)
```




# Load the data

In this section, I retrieve the historical price data for the following 5 stocks:

- **Intesa Sanpaolo S.p.A.**: `ISP.MI` 
- **UniCredit S.p.A.**: `UCG.MI` 
- **Eni S.p.A.**: `ENI.MI` 
- **Stellantis N.V.**: `STLAM.MI` 
- **STMicroelectronics N.V.**: `STMMI.MI` 

The data covers the period from January 1 2019 to Aprile 1 2025.

```{r, include=FALSE, eval=FALSE}
# Get current date
current_date = "2025-04-01"

# Define tickers
tickers = c("ISP.MI", "UCG.MI", "ENI.MI", "STLAM.MI", "STMMI.MI")

# Retrieve adjusted prices for each ticker
# VIX_p = Ad(getSymbols("^VIX", from="2019-01-01", to="2025-02-06", auto.assign=FALSE))
ISP_p = Ad(getSymbols("ISP.MI", from="2019-01-01", to=current_date, auto.assign=FALSE))
UCG_p = Ad(getSymbols("UCG.MI", from="2019-01-01", to=current_date, auto.assign=FALSE))
ENI_p = Ad(getSymbols("ENI.MI", from="2019-01-01", to=current_date, auto.assign=FALSE))
STLAM_p = Ad(getSymbols("STLAM.MI", from="2019-01-01", to=current_date, auto.assign=FALSE))
STMMI_p = Ad(getSymbols("STMMI.MI", from="2019-01-01", to=current_date, auto.assign=FALSE))

# Combine all together
stocks = data.frame(Date = index(ISP_p), ISP_p, UCG_p, ENI_p, STLAM_p, STMMI_p)
colnames(stocks) = c("Date", "ISP", "UCG", "ENI", "STLAM", "STMMI")

# Save the dataframe to the workspace
save(stocks, file="stocks.RData")
```

```{r, echo=FALSE}
load("stocks.RData")
stocks_xts = xts(stocks[,2:6], order.by = stocks$Date)
# compute log_returns
returns = apply(stocks[, -1], 2, function(x) diff(log(x)))
returns = xts(returns, order.by = stocks$Date[-1])
```



# Exploratory Data Analysis

Separate the in-sample data

```{r, include=FALSE}
t_0 = as.Date("2025-01-31")

# For plots and summary statistics
stocks = stocks %>% filter(Date <= t_0)
```

```{r, echo=FALSE}
# Plot the log_10 prices

# Reshape the df to long format
stocks_long = melt(stocks, id.vars = "Date", variable.name = "Stock", value.name = "Price")

# Create list for individual plots
price_plots = list()
stock_names = unique(stocks_long$Stock)

# Loop and create each plot with stock name as title
for (stock in stock_names) {
    stock_data <- subset(stocks_long, Stock == stock)
  
    p <- ggplot(stock_data, aes(x = Date, y = Price)) +
        geom_line(linewidth = 0.8, color = "blue") +
        scale_x_date(date_breaks = "1 year", date_labels = "%Y", date_minor_breaks = "1 month") +
        scale_y_log10() +
        xlab(NULL) + ylab(NULL) +
        ggtitle(stock) +  # Stock name as title
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5))
  
    price_plots[[stock]] <- p
}

# Arrange all plots in a 2x3 grid with a general title
grid.arrange(grobs = price_plots, ncol = 3, nrow = 2)    # , top = "Log Prices"

# Clean up
rm(stocks_long, stock_data, p, price_plots)

```

Now I compute the daily log-returns for each time series.

```{r, echo=FALSE}
# Compute log-returns for each stock and create a df
log_returns = apply(stocks[, -1], 2, function(x) diff(log(x)))
# Multiply the log returns by 100
log_returns = log_returns * 100
log_returns = data.frame(Date = stocks$Date[-1], log_returns)
# tail(log_returns)
```

```{r, echo=FALSE}
# Plot the log-returns

# Reshape the df to long format
log_returns_long = melt(log_returns, id.vars = "Date", variable.name = "Stock", value.name = "Return")

# Create a list to store individual plots
return_plots = list()
stock_names = unique(log_returns_long$Stock)

# Loop through stocks and create plots for log-returns
for (stock in stock_names) {
    stock_data <- subset(log_returns_long, Stock == stock)
  
    # Create the plot with stock name as title
    p <- ggplot(stock_data, aes(x = Date, y = Return)) +
        geom_line(linewidth = 0.5, color = "blue", show.legend = FALSE) +
        scale_x_date(date_breaks = "1 year", date_labels = "%Y", date_minor_breaks = "1 month") +
        ggtitle(stock) +  # Stock name as the title
        xlab(NULL) + ylab(NULL) +  # Remove axis labels
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5))
  
    return_plots[[stock]] <- p  
}

# Arrange all plots in a 2x3 grid with a general title
grid.arrange(grobs = return_plots, ncol = 3, nrow = 2)    # , top = "Log Returns"

# Clean up
rm(stock_data, p, return_plots)

```

```{r, echo=FALSE}
# Plot the histogram of the log-returns

# Create a list to store individual histograms
return_hists = list()

# Loop through stock names to create histograms for log-returns
for (stock in stock_names) {
    stock_data <- subset(log_returns_long, Stock == stock)
  
    # Create the histogram with stock name as title
    p <- ggplot(stock_data, aes(x = Return)) +
        geom_histogram(aes(y = after_stat(density)), bins = 50, fill = "blue", alpha = 0.3, col = "gray31") +
        ggtitle(stock) +  # Stock name as the title
        xlab(NULL) + ylab(NULL) +
        theme_minimal() +
        theme(  
            # axis.text.y = element_blank(),
            plot.title = element_text(hjust = 0.5)
        )
  
    return_hists[[stock]] <- p
}

# Plot in a 2x3 grid with a general title
grid.arrange(grobs = return_hists, ncol = 3, nrow = 2)

# Clean up
rm(log_returns_long, return_hists, stock_data, p)
```

The following dataframe presents key statistical measures and normality test results:  

- **α Trimmed Mean**: the mean calculated after trimming `α` proportion of extreme values.  
- **MAD (Median Absolute Deviation)**: a robust measure of statistical dispersion.  
- **Jarque-Bera Test (p-value)**: tests whether the data follows a normal distribution based on skewness and kurtosis.  
- **Shapiro-Wilk Test (p-value)**: A normality test assessing whether the sample comes from a normal distribution.  

```{r, echo=FALSE}
# Initialize an empty dataframe
stats_df = data.frame(
    Stock = colnames(log_returns[,-1]),
    Trimmed_Mean = numeric(5),
    MAD = numeric(5),
    Jarque_Bera_p = numeric(5),
    Shapiro_Wilk_p = numeric(5)
)

# Compute statistics for each stock
for (i in 1:5) {
    stock_returns = log_returns[,i+1]
  
    stats_df$Trimmed_Mean[i] = round(mean(stock_returns, trim=0.10), 4)
    stats_df$MAD[i] = round(mad(stock_returns), 4)
  
    # Jarque-Bera test
    jb_test = jarque.bera.test(stock_returns)
    stats_df$Jarque_Bera_p[i] = jb_test$p.value  
  
    # Shapiro-Wilk test
    shapiro_test = shapiro.test(stock_returns)
    stats_df$Shapiro_Wilk_p[i] = round(shapiro_test$p.value, 4)
}

print(stats_df)

# Remove variables no longer needed
rm(stock_returns, jb_test, shapiro_test, stats_df)
```

Let us investigate the temporal dependencies and structure within the log-return series.

```{r, echo=FALSE}
# ACF of the log-returns

# Create a list to store individual ACF plots
acf_plots = list()

for (col_name in stock_names) {
    
    series = log_returns[[col_name]]    # Take the log-returns
    n = length(series)  # Number of observations

    # Compute ACF
    series_acf = acf(series, plot = FALSE, lag.max = 20)  # Limit to 20 lags

    # Convert ACF results to a data frame
    acf_df = data.frame(Lag = series_acf$lag, ACF = series_acf$acf)

    # 95% Confidence Intervals for ACF
    conf_int = 1.96 / sqrt(n)

    # Create ACF plot with light blue confidence bands
    p1 = ggplot(acf_df, aes(x = Lag, y = ACF)) +
        geom_ribbon(aes(ymin = -conf_int, ymax = conf_int), fill = "blue", alpha = 0.5) +  # Blue shaded CI
        geom_segment(aes(xend = Lag, yend = 0)) +  # Bars for ACF
        geom_point(color = "blue", size = 2) +     # Points for each lag
        geom_hline(yintercept = 0, linetype = "dashed") +
        geom_hline(yintercept = conf_int, linetype = "dotted", color = "red") +  # Upper CI
        geom_hline(yintercept = -conf_int, linetype = "dotted", color = "red") + # Lower CI
        ggtitle(col_name) +   # Stock name as title
        xlab("Lag") + ylab(NULL) +
        scale_x_continuous(
            breaks = seq(0, 20, by = 1),    # Limit x-axis to 20 lags
            labels = function(x) ifelse(x %% 5 == 0, as.character(x), "")  # Show labels every 5 lags
        ) + 
        scale_y_continuous(limits = c(-0.15, 1)) +
        theme_minimal() +      # Clean theme
        theme(
            # axis.text.y = element_blank(),  # Remove y-axis text for clarity
            plot.title = element_text(hjust = 0.5),  # Center the title
            axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for better visibility
        ) 

    # Store the plot in the list
    acf_plots[[col_name]] = p1
}

# Arrange all ACF plots in a grid (e.g., 2x3 grid)
grid.arrange(grobs = acf_plots, ncol = 3, nrow = 2)

# Clean up
rm(series, n, series_acf, acf_df, conf_int, p1, col_name, acf_plots)

```

Let us examine the temporal dependencies and structure within the absolute value 
of the log-return series.

```{r, echo=FALSE}
# ACF of the absolute value of the log-returns

# Create a list to store individual ACF plots
acf_plots = list()

for (col_name in stock_names) {
    
    series = abs(log_returns[[col_name]])  # Take the absolute log-returns
    n = length(series)  # Number of observations

    # Compute ACF
    series_acf = acf(series, plot = FALSE, lag.max = 20)  # Limit to 20 lags

    # Convert ACF results to a data frame
    acf_df = data.frame(Lag = series_acf$lag, ACF = series_acf$acf)

    # 95% Confidence Intervals for ACF
    conf_int = 1.96 / sqrt(n)

    # Create ACF plot with light blue confidence bands
    p1 = ggplot(acf_df, aes(x = Lag, y = ACF)) +
        geom_ribbon(aes(ymin = -conf_int, ymax = conf_int), fill = "blue", alpha = 0.5) +  # Blue shaded CI
        geom_segment(aes(xend = Lag, yend = 0)) +  # Bars for ACF
        geom_point(color = "blue", size = 2) +     # Points for each lag
        geom_hline(yintercept = 0, linetype = "dashed") +
        geom_hline(yintercept = conf_int, linetype = "dotted", color = "red") +  # Upper CI
        geom_hline(yintercept = -conf_int, linetype = "dotted", color = "red") + # Lower CI
        ggtitle(col_name) +   # Stock name as title
        xlab("Lag") + ylab(NULL) +
        scale_x_continuous(
            breaks = seq(0, 20, by = 1),    # Limit x-axis to 20 lags
            labels = function(x) ifelse(x %% 5 == 0, as.character(x), "")  # Show labels every 5 lags
        ) + 
        scale_y_continuous(limits = c(-0.15, 1)) +
        theme_minimal() +      # Clean theme
        theme(
            # axis.text.y = element_blank(),  # Remove y-axis text for clarity
            plot.title = element_text(hjust = 0.5),  # Center the title
            axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for better visibility
        ) 

    # Store the plot in the list
    acf_plots[[col_name]] = p1
}

# Arrange all ACF plots in a grid (e.g., 2x3 grid)
grid.arrange(grobs = acf_plots, ncol = 3, nrow = 2)

# Clean up
rm(series, n, series_acf, acf_df, conf_int, p1, col_name, acf_plots, log_returns)

```

From the previous plots, we observe the absence of linear dependence in the ACF 
and PACF of the log returns. In contrast, temporal dependencies are evident in 
the corresponding plots of the absolute value of the returns.




# Statistical Analysis

```{r, include=FALSE}
# Define a function that returns the coefficients for the ARMA model
ARMA_coef = function(x) {
    # for seeing all the combinations -> # trace = T
    model = auto.arima(x, max.p = 1, max.q = 1, d = 0, 
                   seasonal = F, ic = "aic")    
    return(as.list(model$coef))
}
```

```{r, include=FALSE}
# Define a function that simulates from the desidered copula model
simulate_from_copula = function(Z, copula_name) {
    
    # set.seed(123)
    
    # Create empty tensor to store simulated standardized residuals
    tensor = array(0, dim = c(L, M, d))
    
    if (copula_name == "skew-t") {
        
        # Fit multivariate skew-t distribution to the standardized residuals
        fit = mst.mple(y = Z, penalty = NULL)

        # Extract estimated parameters from the fitted model
        xi_hat      = as.vector(fit$dp$beta)     # Location vector (ξ) 
        Omega_hat   = fit$dp$Omega               # Scale matrix (Ω)  
        omega_vec   = sqrt(diag(Omega_hat))
        alpha_hat   = fit$dp$alpha               # Skewness vector (α) 
        nu_hat      = fit$dp$nu                  # Degrees of freedom (ν) 
        
        # For each column of the tensor
        for (m in 1:M) {
      
            # Step 1: Simulate L values from the multivariate skew-t distribution
            mst_sim = rmst(n = L, xi = xi_hat, Omega_hat, alpha_hat, nu = nu_hat)
      
            # Step 2: Transform to pseudo-observations via skew-t CDF
            U_new = matrix(0, nrow = L, ncol = d)
            for (i in 1:d) {
                U_new[, i] = sapply(mst_sim[, i], function(y)
                                        pst(y,
                                            xi    = xi_hat[i],
                                            omega = omega_vec[i],
                                            alpha = alpha_hat[i],
                                            nu    = nu_hat
                                        )
                                )
            }
    
            # Step 3: Bound values away from 0 and 1
            U_new = pmin(pmax(U_new, 1e-5), 1 - 1e-5)
            
            # Step 4: Store U_new
            tensor[, m, ] = U_new
        }
    }
    
    
    if (copula_name == "skew-n") {
        
        # Fit multivariate skew-normal distribution to the standardized residuals
        fit = msn.mle(y = Z)
        
        # Extract estimated parameters from the fitted model
        xi_hat      = as.vector(fit$dp$beta)     # Location vector (ξ) 
        Omega_hat   = fit$dp$Omega               # Scale matrix (Ω)  
        omega_vec   = sqrt(diag(Omega_hat))
        alpha_hat   = fit$dp$alpha               # Shape vector (α) 
        
        # For each column of the tensor
        for (m in 1:M) {
      
            # Step 1: Simulate L values from the skew-t copula
            msn_sim = rmsn(n = L, xi = xi_hat, Omega_hat, alpha_hat)
      
            # Step 2: Transform to pseudo-observations via skew-normal CDF
            U_new = matrix(0, nrow = L, ncol = d)
            for (i in 1:d) {
                U_new[, i] = sapply(msn_sim[, i], function(y)
                                            psn(y,
                                                xi    = xi_hat[i],
                                                omega = omega_vec[i],
                                                alpha = alpha_hat[i]
                                            )
                                )
            }
    
            # Step 3: Bound values away from 0 and 1
            U_new = pmin(pmax(U_new, 1e-5), 1 - 1e-5)
            
            # Step 4: Store U_new
            tensor[, m, ] = U_new
        }
    }
    
    return(list(tensor = tensor, alpha = alpha_hat))
}
```

```{r, include=FALSE}
# Computes the optimal portfolio weights by minimizing Entropic Value at Risk (EVaR)
# https://portfoliooptimizationbook.com (the function can be found here)
portfolioEVaR_CVXR = function(X, r_pred, lb, ub, lmd = 1000, alpha = 0.95) {
    T = nrow(X)
    N = ncol(X)
    X = as.matrix(X)
    mu = r_pred
  
    # CVXR
    w = Variable(N)
    s = Variable(1)
    t = Variable(1)
    u = Variable(T)
  
    prob = Problem(Maximize( t(w) %*% mu - lmd*( s - t*log((1-alpha)*T) ) ),
                constraints = list(CVXR:::ExpCone(- X %*% w - s, t*rep(1,T), u),
                                    t >= sum(u),
                                    w >= lb,   
                                    sum(w) == 1,    # full-investing
                                    w <= ub       
                                   )
                )
  
    # Try solving once
    result = tryCatch({
        solve(prob, num_iter = 200, verbose = FALSE)
    }, error = function(e) {
        warning(paste("Solver failed:", conditionMessage(e)))
        return(NULL)
    })

    # If result is invalid or extraction fails, return NULL
    if (is.null(result) || is.null(tryCatch(result$getValue(w), error = function(e) NULL))) {
        return(NULL)
    }
    
    return(as.vector(result$getValue(w)))
}
```

```{r, echo=FALSE}
# Define variables L, M and T (TT)
# The definition of the variables is in the paper presented in the slides in the introduction

stocks_index = index(stocks_xts)
# Number of times I simulate from the copula model
M = 30           # This heavily affects the time of execution
d = length(stock_names)

# L: Length of in-sample window (between 2022-01-01 and t_0)
L = sum(stocks_index > "2022-01-01" & stocks_index <= t_0)
in_sample_dates = index(returns["2022-01-01/"])

# TT: Number of out-of-sample points
TT = sum(stocks_index > t_0) 
# List of dates: t_0 + out-of-sample dates
out_sample_dates = stocks_index[stocks_index >= t_0]

```

```{r}
# Don't run directly all the chunk
# In step 4, select the name of the copula model
# In step 6, change the name of the portfolio weights
# - w_t : skew-t copula portfolio
# - w_n : skew-normal copula portfolio
# Run one time with w_t and one time with w_n

set.seed(1234)

# Vector of portfolio weights
w_t = matrix(0, nrow=TT, ncol=d)            # skew-t copula
w_n = matrix(0, nrow=TT, ncol=d)            # skew-n copula

# Matrix for storing the values of the parameter alpha of the copula
alpha_t = matrix(0, nrow=TT, ncol=d)
alpha_n = matrix(0, nrow=TT, ncol=d)

# First window [t0 - L + 1, t0]
for (j in 1:TT) {         
    
    # Step 1: Subset the dataset
    start_date = in_sample_dates[j]
    end_date = out_sample_dates[j]
    train = returns[paste(start_date, end_date, sep = "/")]
    
    
    
    # Step 2: Model specification and estimation: ARMA + GJR-GARCH(1,1)
    
    # List to store the fitted models
    arma_garch_models = list()

    # Matrix to store standardized residuals 
    Z = matrix(nrow = L, ncol = d)

    # Array to store mu_t_hat and sigma_t_hat
    mu_hat = rep(0, d)
    sigma_hat = rep(0, d)

    for (i in 1:d) {
        
        # Select ARMA parameters
        coefs = ARMA_coef(train[,i])       
        m = ifelse(is.null(coefs$intercept), FALSE, TRUE)
        p = ifelse(is.null(coefs$ar1), 0, 1)  
        q = ifelse(is.null(coefs$ma1), 0, 1) 
        
        # Specify the ARMA-GJR-GARCH model
        gar_spec = ugarchspec(mean.model = list(armaOrder = c(p,q), include.mean = m),
                        variance.model = list(model = "gjrGARCH", garchOrder = c(1,1)),
                        distribution.model = "sstd"    
                      )
         
        # Fit the model
        x = ugarchfit(data=train[,i, drop=F], spec=gar_spec)   # replace train
        arma_garch_models[[i]] = x
        
        # Store standardized residuals
        Z[,i] = as.numeric(residuals(x, standardize=T))
        
        # Forecast conditional mean and volatility
        forecast = ugarchforecast(x, n.ahead = 1)
        mu_hat[i] = as.numeric(forecast@forecast$seriesFor)
        sigma_hat[i] = sqrt(as.numeric(sigma(forecast)))
        
    }

    rm(coefs, m, p, q, gar_spec, forecast)
    
    
    
    # Step 3: Fit semiparametric distribution (SPD) models to marginal standardized residuals
    
    # List for fitted SPD models
    spd_models = list()
    
    for (i in 1:d) {
        # Fit SPD model to the standardized residuals
        x = spdfit(Z[,i])
        # Store the model
        spd_models[[i]] = x
    }
    
    rm(x)
    
    
    
    # Step 4: Simulate from the copula and compute new standardized residuals
    
    # Simulate from fitted copula 
    sfc = simulate_from_copula(Z, "skew-n")
    tensor = sfc$tensor         # all(tensor > 0 & tensor < 1)
    alpha_n[j, ] = sfc$alpha

    # all(tensor >= 0 & tensor <= 1)

    # Transform to standardized residuals using inverse SPD
    for (m in 1:M) {
        for (i in 1:d) {
            x_spd = spd_models[[i]]
            tensor[, m, i] = qspd(tensor[, m, i], x_spd)  
        }
    }

    rm(m, x_spd)
    
    
    
    # Step 5: Predict one-step ahead returns
    
    # Predict standardized residuals as median across simulations
    z_hat = apply(tensor[L, , ], MARGIN = 2, FUN = median)
    
    # One-day ahead forecast of returns
    r_pred = rep(0,d)
    r_pred = mu_hat + z_hat * sigma_hat
    
    
    
    # Step 6: Calculate portfolio weights
    w_new = portfolioEVaR_CVXR(train, r_pred, 0.10, 0.25)

    if (!is.null(w_new)) {
        w_n[j, ] = w_new
    } else {
        w_n[j, ] = w_n[j - 1, ]  # Reuse previous weight vector
    }

    print(j)
    
}
```

```{r, include=FALSE}
# Rebalance portfolio weights every k days (starting from Day 1)
#
# @param  rebalance_every  Must be >=1 and <= number of out-of-sample dates 

rebalance_weights = function(weights_mat, rebalance_every) {
    
    # If rebalance_every = 1, return the original matrix
    if (rebalance_every == 1) return(weights_mat)
    
    n = nrow(weights_mat)
    d = ncol(weights_mat)
    weights_reb = matrix(0, nrow = n, ncol = d)

    # Initialize with Day 1 weights
    current_weights = weights_mat[1, ]  

    for (day in 1:n) {
        if ((day - 1) %% rebalance_every == 0) {
            # update weights on rebalance day
            current_weights = weights_mat[day, ]  
        }
        # assign current weights
        weights_reb[day, ] = current_weights  
    }

    rownames(weights_reb) = rownames(weights_mat)
    return(weights_reb)
}

```

```{r, echo=FALSE}
# https://rpubs.com/DonnieDarkowitz/702869
# out of sample returns
fut_returns = returns[index(returns) > t_0]

# Cumulative return of the equal weight portfolio - Portfolio1
w_EWP = rep(1/d, d)
EWP_returns = fut_returns %*% w_EWP
cum_ret_portf1 = cumprod(1 + EWP_returns) - 1
cum_ret_portf1 = xts(c(0, cum_ret_portf1), order.by = out_sample_dates)

# Cumulative return of portfolio skew-t - Portfolio 2
## Run the next line if you want to rebalance the portfolio
w_t_reb = rebalance_weights(w_t, 5)
portf2_returns = rowSums(w_t_reb * fut_returns)
cum_ret_portf2 = cumprod(1 + portf2_returns) - 1
cum_ret_portf2 = xts(c(0, cum_ret_portf2), order.by = out_sample_dates)

# Cumulative return of portfolio skew-n - Portfolio 3
## Run the next line if you want to rebalance the portfolio
w_n_reb = rebalance_weights(w_n, 5)
portf3_returns = rowSums(w_n_reb * fut_returns)
cum_ret_portf3 = cumprod(1 + portf3_returns) - 1
cum_ret_portf3 = xts(c(0, cum_ret_portf3), order.by = out_sample_dates)

# tail(cbind(EWP_returns, portf2_returns, portf3_returns) * 100, n=10)
round(tail(cbind(cum_ret_portf1, cum_ret_portf2, cum_ret_portf3) * 100, n=1), 4)
```

```{r, include=FALSE, eval=FALSE}
# Weights reported in the thesis in Section 3.4 (table 3.3)
rebalance_indices = seq(1, nrow(w_n_reb), by = 5)
w_n1 = w_n_reb[rebalance_indices, ]
w_t2 = w_t_reb[rebalance_indices, ]
colnames(w_n1) = stock_names
colnames(w_t2) = stock_names
cbind(w_n1, w_t2)
```

```{r, echo=FALSE}
# Combine both cumulative returns into one data frame for easy plotting
cum_returns_df = data.frame(
    Date = index(cum_ret_portf1),
    EWP = coredata(cum_ret_portf1),
    t_copula = coredata(cum_ret_portf2),  
    n_copula = coredata(cum_ret_portf3)  
)

# Melt the data frame to long format
cum_returns_long = melt(cum_returns_df, id.vars = "Date", 
                        variable.name = "Portfolio", value.name = "CumulativeReturn")

# Plotting the cumulative returns
ggplot(cum_returns_long, aes(x = Date, y = CumulativeReturn, color = Portfolio)) +
    geom_line(linewidth = 1) +
    labs(
        title = "Cumulative Returns of EWP vs Copula Portfolios",
        x = "Date", y = "Cumulative Return"
    ) +
    theme_minimal() +
    scale_color_manual(
        values = c("EWP" = "blue", "t_copula" = "red", "n_copula" = "yellow"),
        labels = c("EWP", "t-Copula", "Normal-Copula")
    ) +
    theme(legend.title = element_blank(), legend.position = "bottom") +
    guides(color = guide_legend(title = "Portfolio"))

rm(cum_returns_long)

# ggsave("cumulative_returns_plot.png", width = 10, height = 6, dpi = 300)

```





